Cell Decomposition of Rank 2 Quiver Grassmannians

### 2017-04-25 Grassmannian_cells.tex
Proof of Lemma 4.4, started with improving 4.5

\end{proof}
If we choose subset $\{i_1,\ldots,i_k\}=\{1,\ldots,k\}$, we denote the cokernel simply by $\tilde P_m^{(k)}$. The short exact sequence $\ses{\tilde P_{m-1,k}}{\tilde P_{m}^{(k-1)}}{\tilde P_{m}^{(k)}}$ induces a map between quiver Grassmannians
$$\Psi:Gr^{\tilde Q}_\tbfe(\tilde P_{m}^{(k-1)})\to\bigsqcup_{\tbff+\tbfg=\tbfe} Gr^{\tilde Q}_\tbff( \tilde P_{m-1,k})\times Gr^{\tilde Q}_\tbfg(\tilde P_m^{(k)}),\,U\mapsto (U\cap \tilde P_{m-1,k},\pi(U))$$
which is sometimes called Caldero-Chapoton map.

\begin{lemma}
  If each $Gr^{\tilde Q}_\tbfe(\tilde P_{m})$ has a cell decomposition into affine spaces, then $Gr^{\tilde Q}_\tbfe(\bigoplus_{j=1}^k\tilde P^{l_j}_{m,i_j})$ has a cell decomposition into affine spaces where $l_j\geq 1$.
\end{lemma}
\begin{proof}The claim follows by induction whence considering the map $$\Psi:Gr^{\tilde Q}_\tbfe(\bigoplus_{j=1}^k\tilde P^{l_j}_{m,i_j})\to\bigsqcup_{\tbff+\tbfg=\tbfe}  Gr^{\tilde Q}_\tbff(\tilde P_{m,i_1})\times Gr^{\tilde Q}_\tbfg(\tilde P^{l_1-1}_{m,i_1}\oplus\bigoplus_{j=1}^k\tilde P^{l_j}_{m,i_j}).$$

@@

\begin{lemma}
  \begin{enumerate}
    \item $\tilde P_m^{(n-1)}\cong\tilde P_{m-1}^{(1)}$
    \item The subrepresentation $\bigoplus_{i=1}^{k-1}\tilde P_{m-1,i}\oplus\bigoplus_{i=1}^k\tilde P_{m-2,i}\subset\bigoplus_{i=1}^k\tilde P_{m-1,i}$ is in $(\tilde P_m^{(k)})^\perp$ and
      \[\Ext(\bigoplus_{i=1}^k\tilde P_{m-1,i},\tilde P_m^{(k)})\cong\Ext(\tilde P_{m-1}^{(k)},\tilde P_m^{(k)})\]
      where $\tilde P_{m-1}^{(k)}$ above denotes the cokernel of the inclusion.
  \end{enumerate}
\end{lemma}
\begin{proof}
The first statement follows by the last section.
\end{proof}
\begin{lemma}\label{AR}
Let $m\geq 3$. Consider the short exact sequence $\ses{\tilde P_m}{\tilde P_{m+1}^{(k)}}{\tilde P_{m+1}^{(k-1)}}$.
There exists a short exact sequence 
\[\ses{\tilde P_{m-1}^{n-1}\oplus \tilde P_{m-2}^{n-k}}{\tilde P_m}{\tau\tilde P_{m+1}^{(k-1)}}\]
where $\tau$ denotes the Auslander-Reiten translate. Here the lifts of $P_{m-1}$ and $P_{m-2}$ have to be chosen appropriately. Moreover, we have $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m)\cong \Ext(\tilde P_{m+1}^{(k-1)},\tau\tilde P_{m+1}^{(k-1)})=\kk$, $\Hom(\tilde P_{m+1}^{(k-1)},\tau\tilde P_{m+1}^{(k-1)})=0$ and $\tau\tilde P_{m+1}^{(k-1)}\in ~^\perp(\tilde P_{m+1}^{(k-1)})$. Finally, we have $\Ext(\tilde P_{m+1}^{(k-1)},V)=0$ for every subrepresentation $V$ of $\tilde P_{m-1}^{n-1}\oplus \tilde P_{m-2}^{n-k}$.
%\footnote{introduce/consider/check indices, we have to mod out the corresponding reps in the covering} 
\end{lemma}
\begin{proof}
Actually, we have $(\mathcal  S \tilde P_m)^t=\tilde P_{m+1}$ where $(\quad)^t$ is the functor induced by turning around the arrows (or taking the transpose representation). This means that it is enough to show the statement for $m=3$ as all representations appearing are not isomorphic to the simple injective representation $S_1$.

The last statement needs to be checked and should follow by construction.

\end{proof}
\begin{lemma}\label{quotient}
 Consider the non-split short exact sequence $\ses{\tilde P_m}{\tilde P_{m+1}^{(k)}}{\tilde P_{m+1}^{(k-1)}}$. If $U\subset \tilde P_m$ such that $\Ext(\tilde P_{m+1}^{(k-1)},U)\neq 0$, then we have $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m/U)= 0$ or the fiber $\Psi^{-1}(U,\tilde P_{m+1}^{(k-1)})$ is empty.
\end{lemma}
\begin{proof}
Every $U\subset \tilde P_m$ gives rise to a commutative diagram
  \[\xymatrix{0\ar[r] &\tilde P_{m-1}^{n-1}\oplus \tilde P_{m-2}^{n-k}\ar[r] &  \tilde P_m\ar[r] & \tau\tilde P_{m+1}^{(k-1)}\ar[r] & 0 \\
  0\ar[r]& V\ar[r]\ar[u] & U\ar[u]\ar[r]& W\ar[r]\ar[u]& 0}\]

>>>>

\end{proof}
If we choose subset $\{i_1,\ldots,i_k\}=\{1,\ldots,k\}$, we denote the cokernel simply by $\tilde P_m^{(k)}$.

\begin{lemma}
  If each $Gr^{\tilde Q}_\tbfe(\tilde P_{m})$ has a cell decomposition into affine spaces, then $Gr^{\tilde Q}_\tbfe(\bigoplus_{j=1}^k\tilde P^{l_j}_{m,i_j})$ has a cell decomposition into affine spaces where $l_j\geq 1$.
\end{lemma}
\begin{proof}The claim follows by induction whence considering the map $$\Psi:Gr^{\tilde Q}_\tbfe(\bigoplus_{j=1}^k\tilde P^{l_j}_{m,i_j})\to\bigsqcup_{\tbff+\tbfg=\tbfe}  Gr^{\tilde Q}_\tbff(\tilde P_{m,i_1})\times Gr^{\tilde Q}_\tbfg(\tilde P^{l_1-1}_{m,i_1}\oplus\bigoplus_{j=1}^k\tilde P^{l_j}_{m,i_j}),\,U\mapsto (U\cap \tilde P_{m-1,i_1},\pi(U))$$
which is also sometimes called Caldero-Chapoton map.

@@

\begin{lemma}\footnote{do we need this?}
  \begin{enumerate}
    \item $\tilde P_m^{(i_1,\ldots,i_{n-1})}\cong\tilde P_{m-1}^{(i_n)}$
    \item The subrepresentation $\bigoplus_{i=1}^{k-1}\tilde P_{m-1,i}\oplus\bigoplus_{i=1}^k\tilde P_{m-2,i}\subset\bigoplus_{i=1}^k\tilde P_{m-1,i}$ is in $(\tilde P_m^{(k)})^\perp$ and
      \[\Ext(\bigoplus_{i=1}^k\tilde P_{m-1,i},\tilde P_m^{(k)})\cong\Ext(\tilde P_{m-1}^{(k)},\tilde P_m^{(k)})\]
      where $\tilde P_{m-1}^{(k)}$ above denotes the cokernel of the inclusion.
  \end{enumerate}
\end{lemma}
\begin{proof}
The first statement follows for instance by induction when applying BGP-reflections.
\end{proof}
\begin{lemma}\label{AR}
Let $m\geq 3$ and $n\geq k\geq 2$. Consider the short exact sequence $\ses{\tilde P_{m,k}}{\tilde P_{m+1}^{(k-1)}}{\tilde P_{m+1}^{(k)}}$. 
\begin{enumerate}
\item There exists a short exact sequence 
\[\ses{\bigoplus_{\substack{1\leq j\leq n\\j\neq k}}\tilde P_{m-1,(k,j)}\oplus \bigoplus_{j=1}^k\tilde P_{m-2,(k,k,j)}}{\tilde P_{m,k}}{\tau\tilde P_{m+1}^{(k)}}\]
where $\tau$ denotes the Auslander-Reiten translate where
$\tau\tilde P_{m+1}^{(k)}=\tilde P_{m-1}^{(k)}$. 
\item Moreover, we have $\Ext(\tilde P_{m+1}^{(k)},\tilde P_{m,k})\cong \Ext(\tilde P_{m+1}^{(k)},\tau\tilde P_{m+1}^{(k)})=\kk$, $\Hom(\tilde P_{m+1}^{(k)},\tau\tilde P_{m+1}^{(k)})=0$. 
\item We have $\tau\tilde P_{m+1}^{(k)}\in ~^\perp(\tilde P_{m+1}^{(k)})$\footnote{do we need this?}. 
%\item Finally, we have $\Ext(\tilde P_{m+1}^{(k-1)},V)=0$ for every subrepresentation $V$ of $\tilde P_{m-1}^{n-1}\oplus \tilde P_{m-2}^{n-k}$.
\end{enumerate}%\footnote{introduce/consider/check indices, we have to mod out the corresponding reps in the covering} 
\end{lemma}
\begin{proof}
Actually, we have $(\mathcal  S \tilde P_m)^t=\tilde P_{m+1}$ where $(\quad)^t$ is the functor induced by turning around the arrows (or taking the transpose representation). This means that it is enough to show the statement for $m=3$ as all representations appearing are not isomorphic to the simple injective representation $S_2$. In this case, the first statement is straightforward (maybe a diagram?!).

It is also straightforward to check that $\bigoplus_{j\neq k}\tilde P_{2,(k,j)}\oplus \bigoplus_{j=1}^k\tilde P_{1,(k,k,j)}\in (\tilde P_{4}^{(k)})^\perp$ and, moreover, that $\Hom(\tilde P_{4}^{(k)},\tilde P_{3,k})=0$ which yields the second claim when considering the long exact sequence obtained when applying $\Hom(\tilde P_{4}^{(k)},\underline{\quad})$.

\end{proof}
The short exact sequence $\ses{\tilde P_{m-1,k}}{\tilde P_{m}^{(k-1)}}{\tilde P_{m}^{(k)}}$ induces a map between quiver Grassmannians
$$\Psi_1:Gr^{\tilde Q}_\tbfe(\tilde P_{m}^{(k-1)})\to\bigsqcup_{\tbff+\tbfg=\tbfe} Gr^{\tilde Q}_\tbff( \tilde P_{m-1,k})\times Gr^{\tilde Q}_\tbfg(\tilde P_m^{(k)}).$$
Let $P(m,k):=\bigoplus_{\substack{1\leq j\leq n\\j\neq k}}\tilde P_{m-1,(k,j)}\oplus \bigoplus_{j=1}^k\tilde P_{m-2,(k,k,j)}$.

By Lemma \ref{AR}, every $U\subset \tilde P_m$ gives rise to a commutative diagram

 \[\xymatrix{0\ar[r] &P(m,k)\ar[r] &  \tilde P_m\ar[r] & \tau\tilde P_{m+1}^{(k-1)}\ar[r] & 0 \\
  0\ar[r]& V\ar[r]\ar[u] & U\ar[u]\ar[r]& W\ar[r]\ar[u]& 0}\]
By $\Psi_2$ we denote the corresponding map of quiver Grassmannians.
Using this notation, we obtain:
\begin{lemma}\label{quotient}
 Consider the non-split short exact sequence $\ses{\tilde P_m}{\tilde P_{m+1}^{(k)}}{\tilde P_{m+1}^{(k-1)}}$. Let $U\subset \tilde P_m$ such that $\Ext(\tilde P_{m+1}^{(k-1)},U)\neq 0$. Then we have
\begin{enumerate}
\item The fibre $\Psi^{-1}(U,\tilde P_{m+1}^{(k-1)})$ is empty the case if and only if $U\in\Psi^{-1}_2(0,W)$ for some $W\neq 0$.
\item If the fibre is not empty, we have $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m/U)= 0$.
\end{enumerate} 
\end{lemma}
\begin{proof}
